---
title: "Data and Cleaning"
output: html_document
---

Data cleaning and manipulation process takes place in this document. First, I begin by loading in my packages.

#Packages
```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(janitor)
```

Next, I load in the SIS data and conduct feature engineering to get a scramble column into our PBP data. I also load in the players total points so we can later control for who ran the route and who threw the pass when determining how effective a route combo actually is. 

#SIS Data
```{r}
#plays
pbp     <- read_csv("https://github.com/SportsInfoSolutions/AnalyticsChallenge2021/raw/main/Data/PlayByPlay.csv") %>%
  #add scrambles
  mutate(scramble = if_else(str_detect(PlayDesc, "scramble"), 1, 0)) %>%
  #if scramble reclassify as a passing play
  mutate(EventType = ifelse(scramble == 1, "pass", EventType)) %>%
  #now we can grab all of our passes
  filter(EventType == "pass" | EventType == "Challenge Pass") %>%
  #and then remove spikes
  filter(Spike == 0 | Spike == "NULL") 

#players
routes  <- read_csv("https://github.com/SportsInfoSolutions/AnalyticsChallenge2021/raw/main/Data/SkillPositionPlayers.csv") %>%
  left_join(pbp, by = c("GameID", "EventID")) %>%
  #remove plays with an NA because those are non passing plays
  filter(is.na(EventType) == FALSE) 

#player ratings on each game, taking the sum of their season performance
players <- read_csv("https://github.com/SportsInfoSolutions/AnalyticsChallenge2021/raw/main/Data/PlayerTotalPoints.csv") %>%
  #remove null games
  filter(TotalPoints != "NULL") %>%
  #make integer
  mutate(TotalPoints = as.integer(TotalPoints)) %>%
  group_by(PlayerId, Name) %>%
  #get season total points
  summarize(Total_Points = sum(TotalPoints)) %>%
  arrange(desc(Total_Points)) %>%
  ungroup()

#will use to determine QB in a play and later control for them
throwers <- routes %>%
  filter(OnFieldPosition == "QB") %>%
  dplyr::select(GameID, EventID, PlayerId, Name)

#will use to add coverages to non targeted plays later
coverages <- routes %>%
  dplyr::select(GameID,
         EventID,
         CoverageScheme) %>%
  distinct()
```

One quick fix to our data is for fumbles. If a play gained 20 yards but was fumbled away then it would be misleading to say the route combo had a negative EPA. Thus, we will add in NFLFastR data to get the value of the play in EPA leading up to the fumble. We then replace fumble play EPA with the proper EPA adjustment.

#Replace lost fumbles with EPA up to point of fumble being lost
```{r}
#NFLFastR data to clean up plays where there are fumbles
pbp_NFLFastR <- read_rds("NFL_Fast_R_2020.rds")

#replace plays where fumble is lost with EPA up to the point of fumble
pbp_clean <- pbp %>%
  left_join(pbp_NFLFastR, by = c("TimeLeft" = "quarter_seconds_remaining",
                                 "DefensiveTeam",
                                 "OffensiveTeam",
                                 "Week" = "week")) %>%
  #replace fumble epa with proper epa
  mutate(EPA = ifelse(FumbleByReceiver == "1", qb_epa, EPA)) 
```

Now I begin to examine my routes, there are about 50 different routes in the data. This needs to be condensed to get more uniformity in our route combos later. If we do not adjust our routes like below then it would be difficult to pull highly used route combos because of very slight differences. 

#clean route data
```{r}
#check route types
length(unique(routes$Route))

#fix routes and generalize a few to avoid too many different route combos
routes_clean <- routes %>%
  #remove these routes which we cannot determine a side to which they were run
  filter(Route != "NULL", Route != "Chip", Route != "Check & Release", Route != "Blocking") %>%
  #fix certain routes for consistency, also removing spaces because of the formatting I conduct later
  mutate(
    Route = case_when(
      Route == "Fade - Back Shoulder" ~ "Go",
      Route == "Fade" ~ "Go",
      Route == "Screen - Bubble" ~ "Screen",
      Route == "Screen - TE" ~ "Screen",
      Route == "Screen - Quick" ~ "Screen",
      Route == "Screen - Quick" ~ "Screen",
      Route == "Screen - Tunnel" ~ "Screen",
      Route == "Screen - Beneath" ~ "Screen",
      Route == "Screen - Shovel" ~ "Screen",
      Route == "Screen - Drag" ~ "Screen",
      Route == "Screen - RB" ~ "Screen",
      Route == "Chip - Flat" ~ "Flat",
      Route == "Chip - Curl" ~ "Curl",
      Route == "Chip - Drag" ~ "Drag",
      Route == "Chip - Seam" ~ "Seam",
      Route == "Stick - Nod" ~ "Seam",
      Route == "Beneath" ~ "Drag",
      Route == "Hitch & Go" ~ "DoubleMove",
      Route == "Out & Up" ~ "DoubleMove",
      Route == "Go/Fly" ~ "Go",
      Route == "Over Ball" ~ "SitOverMiddle",
      Route == "Corner Post" ~ "COPO", #short for corner post
      Route == "Post Corner" ~ "POCO", #short for post corner
      Route == "Deep Cross" ~ "DeepCross",
      Route == "Jet Sweep Pass" ~ "JetSweepPass",
      TRUE ~ Route
      )) 

#check our unique routes again, grouped down to about 30 and it will go down with the next chunk where we fix flats and swings
unique(routes_clean$Route)
```

Now that routes are condensed, I will continue to condense a little more. In this case, with routes that are marked as going to the right or left but have an opposite side of center. I think there is an issue with the data where individuals who are on the right but have a route such as flat left with no motion should actually be something like a drag route. Also, for the sake of getting some condensed data I am moving swings designated to one side but lined up on the other as the innermost receiver on the route run side. 

#Fix routes that state side to which they are run so that they are marked as part of the proper route combo 
```{r}
routes_clean <- routes_clean %>%
  
  #mutate so routes that state flat to one side of field are renamed as just a drag route
  mutate(Route = ifelse(Route == "Flat - Right" & FastMotion == 0 & SideOfCenter == "L", "Drag",  Route)) %>%
  mutate(Route = ifelse(Route == "Flat - Left"  & FastMotion == 0 & SideOfCenter == "R", "Drag",  Route)) %>%

  #make sure any swings that release left or right are classified as a side of center (ignoring motion for a moment) & proper order
  mutate(Order_OutsideToInside = ifelse(Route == "Swing - Right" & FastMotion== 0& SideOfCenter== "L","Fix",Order_OutsideToInside)) %>%
  mutate(Order_OutsideToInside = ifelse(Route == "Swing - Left"  & FastMotion== 0& SideOfCenter== "R","Fix",Order_OutsideToInside)) %>%
  mutate(SideOfCenter = ifelse(Route == "Swing - Right"& FastMotion == 0 & SideOfCenter == "L", "R", SideOfCenter)) %>%
  mutate(SideOfCenter = ifelse(Route == "Swing - Left" & FastMotion == 0 & SideOfCenter == "R", "L", SideOfCenter)) %>%
  
  #pickup the fast motion flats and swings that start on other side of field and make their side opposite and order 5 (will fix later)
  mutate(Order_OutsideToInside = ifelse(Route == "Flat - Right" &FastMotion== 1 & SideOfCenter== "L","Fix", Order_OutsideToInside)) %>%
  mutate(Order_OutsideToInside = ifelse(Route == "Flat - Left"  &FastMotion== 1 & SideOfCenter== "R","Fix", Order_OutsideToInside)) %>%
  mutate(Order_OutsideToInside = ifelse(Route == "Swing - Right"&FastMotion== 1 & SideOfCenter== "L","Fix", Order_OutsideToInside)) %>%
  mutate(Order_OutsideToInside = ifelse(Route == "Swing - Left" & FastMotion == 1&SideOfCenter =="R","Fix", Order_OutsideToInside)) %>%
  
  #change the fast motion flats and swings to the proper side now
  mutate(SideOfCenter = ifelse(Route == "Flat - Right" , "R", SideOfCenter)) %>%
  mutate(SideOfCenter = ifelse(Route == "Flat - Left"  , "L", SideOfCenter)) %>%
  mutate(SideOfCenter = ifelse(Route == "Swing - Right", "R", SideOfCenter)) %>%
  mutate(SideOfCenter = ifelse(Route == "Swing - Left" , "L", SideOfCenter)) %>%
  
  #fix running back routes with the flat or swing description who are missing an order
  mutate(Order_OutsideToInside = ifelse(Route == "Flat - Right" & Order_OutsideToInside == "NULL", "Fix", Order_OutsideToInside)) %>%
  mutate(Order_OutsideToInside = ifelse(Route == "Flat - Left"  & Order_OutsideToInside == "NULL", "Fix", Order_OutsideToInside)) %>%
  mutate(Order_OutsideToInside = ifelse(Route == "Swing - Right"& Order_OutsideToInside == "NULL", "Fix", Order_OutsideToInside)) %>%
  mutate(Order_OutsideToInside = ifelse(Route == "Swing - Left" & Order_OutsideToInside == "NULL", "Fix", Order_OutsideToInside)) %>%

  #make so any flats & swings that release left or right are classified as just a flat or swing now 
  mutate(Route = ifelse(Route == "Flat - Right" , "Flat",  Route)) %>%
  mutate(Route = ifelse(Route == "Flat - Left"  , "Flat",  Route)) %>%
  mutate(Route = ifelse(Route == "Swing - Right", "Swing", Route)) %>%
  mutate(Route = ifelse(Route == "Swing - Left" , "Swing", Route)) %>%
  
  #remove routes that we cannot infer side of center from
  filter(SideOfCenter != "NULL")

#check our new distribution of routes, much better
unique(routes_clean$Route)

```

One task before advancing is to pull who was targeted on each play so we use that in our analysis of targeted route combos.

#Get info on route combo that was targeted and combine with total points data and game data
```{r}
Target_Info <- routes_clean %>%
  filter(Target == 1) %>%
  dplyr::select(GameID, EventID, PlayerId, Name, SideOfCenter, Target, Route, OnFieldPosition) %>%
  mutate(Side_Targeted = SideOfCenter) %>%
  dplyr::select(-SideOfCenter) %>%
  #add in pass catchers
  left_join(players, by = c("PlayerId", "Name")) %>%
  rename("Receiver" = "Name") %>%
  rename("WR_Points" = "Total_Points") %>%
  dplyr::select(-PlayerId) %>%
  #add in who was the QB 
  left_join(throwers, by = c("GameID", "EventID")) %>%
  #read in players data for QBs 
  left_join(players, by = c("PlayerId", "Name")) %>%
  rename("QB" = "Name") %>%
  rename("QB_Points" = "Total_Points") 
```

Similar to above, a factor that we may want to control for is if the play was play action so we will pull that out for later analysis.

#pull out the play action plays and add them to our plays
```{r}
#find plays with play action
PA_Plays <- routes_clean %>%
  filter(Route == "Run Fake") %>%
  dplyr::select(GameID,
         EventID) %>%
  mutate(Play_Action = 1)

#merge in PA to our plays
pbp_clean <- pbp_clean %>%
  left_join(PA_Plays, by = c("GameID", "EventID")) %>%
  mutate(Play_Action = ifelse(is.na(Play_Action) == TRUE, 0 ,1))
```

Now we have two tasks to take care of, 1.) fix the ordering of routes because we removed certain routes during cleaning and added routes to the opposite side of the field and 2.) we want to obtain route interactions, meaning routes that crossed the field and are part of the opposite field route combo. First, we will fix the ordering

#fix ordering for the mutations we did with running backs and motions
```{r}
#make numeric which replaces our "fix" designation with an NA
routes_clean$Order_OutsideToInside <- as.numeric(routes_clean$Order_OutsideToInside)

#get max and min order to ID bad rows
routes_clean <- routes_clean %>%
  group_by(GameID, 
           EventID,
           SideOfCenter) %>%
  #create var that identifies how many players are on one side of a formation, if max and min is NA we need to fix
  mutate(max_route         = max(Order_OutsideToInside, na.rm = TRUE),
         min_route         = min(Order_OutsideToInside, na.rm = TRUE),
         number_of_players = n()) %>%
  mutate() %>%
  ungroup() %>%
  mutate(max_route = ifelse(max_route == "-Inf", NA, max_route),
         min_route = ifelse(max_route == "-Inf", NA, min_route)) %>%
  #make it so the player is classified as 1 if there are no other routes on their side of the field
  group_by(GameID, 
           EventID,
           SideOfCenter) %>%
  #statement that says if nobody else is on side of the route combo then make order 1
  mutate(Order_OutsideToInside = ifelse(is.na(max_route) == TRUE & is.na(min_route) == TRUE, 1, Order_OutsideToInside)) %>%
  ungroup() %>%
  #make any added routes plus 1 to the max route
  mutate(Order_OutsideToInside = ifelse(is.na(Order_OutsideToInside) == TRUE, max_route +1, Order_OutsideToInside))


#check to make sure our order is good and it is!
routes_clean %>%
  filter(is.na(max_route) == TRUE & is.na(min_route) == TRUE)

#now we need to move routes up a slot if there is not any number 1 in the order
routes_clean <- routes_clean %>%
    group_by(GameID, 
           EventID,
           SideOfCenter) %>%
  mutate(min_route = min(Order_OutsideToInside, na.rm = TRUE)) %>%
  mutate(Order_OutsideToInside = ifelse(min_route > 1, Order_OutsideToInside-1, Order_OutsideToInside)) %>%
  #same order that says if there is no number 1 then move each wr over 
  group_by(GameID, 
           EventID,
           SideOfCenter) %>%
  mutate(min_route = min(Order_OutsideToInside, na.rm = TRUE)) %>%
  mutate(Order_OutsideToInside = ifelse(min_route > 1, Order_OutsideToInside-1, Order_OutsideToInside)) %>%
  #same order that says if there is no number 1 then move each wr over 
  group_by(GameID, 
           EventID,
           SideOfCenter) %>%
  mutate(min_route = min(Order_OutsideToInside, na.rm = TRUE)) %>%
  mutate(Order_OutsideToInside = ifelse(min_route > 1, Order_OutsideToInside-1, Order_OutsideToInside)) %>%
  #same order that says if there is no number 1 then move each wr over 
  group_by(GameID, 
           EventID,
           SideOfCenter) %>%
    mutate(min_route = min(Order_OutsideToInside, na.rm = TRUE)) %>%
  mutate(Order_OutsideToInside = ifelse(min_route > 1, Order_OutsideToInside-1, Order_OutsideToInside)) %>%
  #same order that says if there is no number 1 then move each wr over 
  group_by(GameID, 
           EventID,
           SideOfCenter) %>%
  mutate(max_route = max(Order_OutsideToInside, na.rm = TRUE))

#one issue that has come from this are plays where there are two players in a swing or flat that we reclassified and they cannot be properly ordered so we will remove those plays
routes_clean <- routes_clean %>%
  group_by(GameID, EventID, SideOfCenter, Order_OutsideToInside) %>%
  summarize(summed = sum(Order_OutsideToInside)) %>%
  mutate(diff = summed - Order_OutsideToInside) %>%
  ungroup() %>%
  dplyr::select(GameID, EventID, diff, summed) %>%
  group_by(GameID, EventID) %>%
  summarize(remove = ifelse(sum(diff) != 0, 1,0)) %>%
  distinct() %>%
  left_join(routes_clean, by = c("GameID", "EventID")) %>%
  filter(remove == 0) %>%
  filter(Route != "Run Fake")

#pivot data, looks good! Now to address our second issue of crossing routes
routes_clean %>%
  pivot_wider(
    id_cols = c(GameID, EventID, SideOfCenter),
    names_from = Order_OutsideToInside,
    values_from = Route
  ) %>%
  filter(`2` >1)
```

Now that our data is cleaned we can address our second issue of having crossing routes that interact with out routes on the field. My methodology here is that if a route is a crosser aligned at the slot or TE position and the other side of the field has a crosser at any spot then we should have an interaction. Otherwise we will ignore. 

#find crossing combos
```{r}
#routes we will classify as crosser
crossing_routes <- c("Drag", "Dig", "Deep Cross", "Post", "Sit Over Middle")

#find plays where crossers from both side of the field
Routes_With_Cross <- routes_clean %>%
  mutate(Crosser = ifelse(OnFieldPosition == "SWR" & str_detect(Route, paste(crossing_routes, collapse = "|")), 1, 
                   ifelse(OnFieldPosition == "TE"  & str_detect(Route, paste(crossing_routes, collapse = "|")), 1, 0))) %>%
  dplyr::select(EventID,
         GameID,
         Route,
         Crosser,
         SideOfCenter,
         Target) %>%
  filter(Crosser == 1) %>%
  group_by(EventID,
           GameID) %>%
  mutate(Crossing_Pattern_Targeted = sum(Target)) %>%
  ungroup() %>%
  dplyr::select(-Target,
         -Crossing_Pattern_Targeted) %>%
  pivot_wider(
    id_cols = c(GameID, EventID),
    names_from = SideOfCenter,
    values_from = Route,
    values_fn = length
  ) %>%
  filter(is.na(`L`) == FALSE & is.na(`R`) == FALSE) %>%
  mutate(Crossing_Combo = 1) %>%
  dplyr::select(GameID,
         EventID,
         Crossing_Combo)

#join data so no we know plays that involve a crossing route combo, now we will save two data frames for easier breakdown
Plays_With_Crossing_Patterns    <- routes_clean %>%
  left_join(Routes_With_Cross, by = c("GameID", "EventID")) %>% 
  mutate(Crossing_Combo = ifelse(is.na(Crossing_Combo) == TRUE, 0, Crossing_Combo)) %>%
  filter(Crossing_Combo == 1) 

#ready for later analysis
Plays_Without_Crossing_Patterns <- routes_clean %>%
  left_join(Routes_With_Cross, by = c("GameID", "EventID")) %>% 
  mutate(Crossing_Combo = ifelse(is.na(Crossing_Combo) == TRUE, 0, Crossing_Combo)) %>%
  filter(Crossing_Combo == 0) %>%
  pivot_wider(
    id_cols = c(GameID, EventID, SideOfCenter),
    names_from = Order_OutsideToInside,
    values_from = Route
  ) %>%
  filter(!is.na(`2`)) %>%
  mutate(
    combo = glue::glue("{`1`} - {`2`} {ifelse(is.na(`3`), '', '-')} {ifelse(!is.na(`3`), `3`, '')}")
  ) %>%
  #join in the target and player data for later, keeping route combos with no targets for the visual of most popular route combos
  left_join(Target_Info, by = c("GameID",
                                "EventID"))  %>%
  dplyr::select(-`1`,
         -`2`,
         -`3`,
         -`4`) %>%
  left_join(coverages, by = c("GameID", "EventID"))
```

So now we have a group of plays with crossing route combos to extract. We will do this below.

#Extract crossing route patterns
```{r}
Plays_To_Merge <- Plays_With_Crossing_Patterns %>%
  #filter for only inside wr's so we do not grab outside WR's by accident
  filter(OnFieldPosition == "SWR" | OnFieldPosition == "TE")%>%
  mutate(Crossing_Route = ifelse(str_detect(Route, paste(crossing_routes, collapse = "|")),1,0)) %>%
  filter(Crossing_Route == 1) %>%
    pivot_wider(
    id_cols = c(GameID, EventID, SideOfCenter),
    names_from = Order_OutsideToInside,
    values_from = Route
  ) %>%
    mutate(
    combo = glue::glue("{ifelse(is.na(`1`), '', '')} {ifelse(!is.na(`1`), `1`, '')} {ifelse(is.na(`2`), '', '')} {ifelse(!is.na(`2`), `2`, '')} {ifelse(is.na(`3`), '', '')} {ifelse(!is.na(`3`), `3`, '')}{ifelse(is.na(`4`), '', '')} {ifelse(!is.na(`4`), `4`, '')}")
  ) %>%
  #delete leading //
  mutate(combo = trimws(combo, which = c("both", "left", "right"), whitespace = "[ \t\r\n]"),
         combo = gsub("\\s+"," ",combo),
         combo = gsub(" ", " - ", combo)) %>%
   dplyr::select(GameID,
         EventID,
         SideOfCenter,
         combo) %>%
  pivot_wider(
    id_cols = c(GameID, EventID),
    names_from = SideOfCenter,
    values_from = combo
    ) %>%
  rowwise() %>%   
  mutate(combo = paste(sort(c(L, R)), collapse = " ----- ")) %>%
  dplyr::select(GameID,
         EventID,
         combo) %>%
  left_join(Target_Info, by = c("GameID",
                                "EventID")) %>%
  mutate(Crosser_Targeted = ifelse(OnFieldPosition == "SWR" & str_detect(Route, paste(crossing_routes, collapse = "|")), 1,
                            ifelse(OnFieldPosition == "TE"  & str_detect(Route, paste(crossing_routes, collapse = "|")), 1, 0))) %>%
  dplyr::select(-Side_Targeted) %>%
  left_join(coverages, by = c("GameID", "EventID")) %>%
  mutate(SideOfCenter = "C")

```

And we have our crossing route combos along with our non crossing route combos! Now we want to merge our crossers and non crossers into one data frame for analysis.

#Merge crossers and non crossers
```{r}
#add in sacks for a later portion
pbp <- read_csv("https://github.com/SportsInfoSolutions/AnalyticsChallenge2021/raw/main/Data/PlayByPlay.csv") %>%
  #add scrambles
  mutate(scramble = if_else(str_detect(PlayDesc, "scramble"), 1, 0)) %>%
  #if scramble reclassify as a passing play
  mutate(EventType = ifelse(scramble == 1, "pass", EventType)) %>%
  #now we can grab all of our passes
  filter(EventType == "pass" | EventType == "Challenge Pass") %>%
  #and then remove spikes
  filter(Spike == 0 | Spike == "NULL") %>%
  rename("EPA_Final" = "EPA")

Plays_Without_Crossing_Patterns %>%
  mutate(Crosser_Targeted = NA) %>%
  rbind(Plays_To_Merge) %>%
  left_join(pbp, by = c("GameID", "EventID")) %>%
  write_rds("Plays_For_Analysis.rds")
```



