---
title: "EPA Model"
output: html_document
---

#Packages
```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(janitor)
library(gamm4)
library(purrr)
library(ggplot2)
library(nflfastR)
library(scales)
library(brms)
library(zoo)
library(stats4)
library(VGAM)
library(gamlss)
library(broom)
library(ffscrapr)
library(rethinking)
```

Similar to the completion percentage model, we can estimate the effect of a route combo on the EPA of a play while controlling for confounding factors. 

#Data for modeling
```{r}
#load all route combos
Data <- read_rds("Plays_For_Analysis.rds") %>%
  ungroup()%>%
  dplyr::select(-CoverageScheme.y) %>%
  rename("CoverageScheme" = "CoverageScheme.x") %>%
  #remove prevent, screen and other defense from our coverage 
  filter(CoverageScheme != "Screen",
         CoverageScheme != "Other",
         CoverageScheme != "Prevent",
         CoverageScheme != "NULL") %>%
  #doing this for filtering purposes
  filter(is.na(EPA_Final) == FALSE) %>%
  mutate(Completion = as.integer(Completion),
         Attempt = as.integer(Attempt),
         ThrowDepth = as.integer(ThrowDepth),
         combo = trimws(combo)) %>%
  mutate(ID_Play = paste0(GameID,EventID),
         Side_Targeted = ifelse(is.na(Side_Targeted) == TRUE, "C", Side_Targeted)) %>%
  #filter for targets only
  filter(Target == 1) %>%
  filter(SideOfCenter == Side_Targeted)%>%
  dplyr::select(combo,
                CoverageScheme,
                EPA_Final,
                Attempt,
                Completion,
                ThrowDepth,
                QB,
                WR_Points,
                QB_Points,
                Down,
                ToGo,
                PressureOnPlay,
                StartYard,
                SideOfField) %>%
  mutate(CoverageScheme = as.factor(CoverageScheme),
         combo_coverage = paste0(combo, " | ", CoverageScheme),
         Combo_Cov_ID = as.numeric(factor(str_c(combo_coverage))),
         Coverage_ID = as.numeric(factor(str_c(CoverageScheme)))) %>%
  #fix a few columns for the completion percentage model
  dplyr::mutate(
    ThrowDepth    = ifelse(ThrowDepth < 0, 0 , ThrowDepth),
    ayard_is_zero = if_else(ThrowDepth == 0, 1, 0),
    Difference    = 50 - StartYard,
    StartYard     = ifelse(SideOfField == "Oppo", StartYard + (Difference*2), StartYard)) %>%
  filter(QB != "Kendall Hinton",
         QB != "Russell Gage",
         QB != "Mecole Hardman") #just no
```

First, we need to adjust Quarterbacks who have a low number of attempts which is creating a non representative EPA for them. 

```{r}
#demonstrate our issue, Kyle Allen should not be estimated as among our best QB's
Data %>%
  group_by(QB) %>%
  summarize(attempts = sum(Attempt),
            epa = mean(EPA_Final)) %>%
  arrange(desc(epa))

#we can take the QB EPA per play throughout their career to get an estimate of their EPA per play
pbp <- read_rds("NFL_Fast_R_CP.rds")

passers <- Data %>%
  mutate(QB = dp_cleannames(QB)) %>%
  group_by(QB) %>%
  summarize(attempts = sum(Attempt),
            epa = mean(EPA_Final)) %>%
  arrange(desc(epa))

Full_Names <- nflfastR::fast_scraper_roster(2020) %>%
  filter(position == "QB") %>%
  dplyr::select(full_name,
                gsis_id) %>%
  arrange(full_name) %>%
  mutate(full_name = ifelse(full_name == "C.J. Beathard", "CJ Beathard", full_name))

#make our adjustment to EPA
Passer_Adjustment <- pbp %>%
  filter(is.na(passer_player_id) == FALSE, is.na(epa) == FALSE) %>%
  group_by(passer_player_id) %>%
  summarize(Career_EPA = mean(epa, na.rm = TRUE),
         Career_Attempts = sum(pass_attempt, na.rm = TRUE)) %>%
  ungroup() %>%
  left_join(Full_Names, by = c("passer_player_id" = "gsis_id")) %>%
  filter(is.na(full_name) == FALSE) %>%
  left_join(passers, by = c("full_name" = "QB")) %>%
  filter(is.na(attempts) == FALSE) %>%
  mutate(Career_EPA = ifelse(Career_Attempts <= 10, -0.05929128, Career_EPA)) %>%
  arrange(desc(Career_EPA)) %>%
  dplyr::select(passer_player_id,
                Career_EPA,
                full_name) %>%
  arrange(full_name)

#add to SIS data
Data <- Data %>%
  mutate(QB = dp_cleannames(QB)) %>%
  left_join(Passer_Adjustment, by = c("QB" = "full_name")) %>%
  mutate(combo = as.factor(combo),
         combo_ID = as.numeric(factor(str_c(combo))))

rm(pbp)
rm(Full_Names)
rm(passers)
rm(Passer_Adjustment)
```

Now we can build the model of EPA and get our estimates. We will use a Bayesian mixed model so we can establish a prior on EPA which has an unusual distribution. A student T is appropriate here to capture outlier events.

#Establish model using student T and fit model.
```{r}
#Bayesian mixed model

EPA_Mixed <- brms::brm(
  EPA_Final ~ Career_EPA + (1 | combo_ID:Coverage_ID),
  data  = Data,
  control = list(adapt_delta = 0.99),
      iter = 3000, warmup = 1000, chains = 3,
      seed = 9)

```

Now we can extract our random effects and obtain our best route combos.

#Get combo codes
```{r}
#get route combos for later merging
combos <- Data %>%
  mutate(id = paste0(combo_ID, "_", Coverage_ID)) %>%
  group_by(id) %>%
  dplyr::summarise(
    combo = unique(combo),
    coverage = unique(CoverageScheme),
    occurances = n()
  ) %>%
  ungroup() %>%
  mutate(id = as.character(id)) %>%
  arrange(desc(occurances))
```

#extract estimates
```{r}
est <- mixedup::extract_random_effects(EPA_Mixed) %>%
  arrange(value) %>%
  dplyr::rename("id" = "group") 

# Prepare data for plot
est <- est %>%
  left_join(combos , by = "id") %>%
  arrange(value) %>%
  filter(is.na(coverage) == FALSE) %>%
  filter(is.na(combo) == FALSE) %>%
  dplyr::select(
    -group_var,
    -effect,
    -id
  ) %>%
  arrange(desc(value))
```

```{r}
est  %>%
  filter(occurances > 9) %>%
  filter(coverage == "Cover 3") %>%
  head(15) %>%
  ggplot(aes(x = reorder(combo, value), value)) +
  geom_linerange(size = .5, aes(
    ymin = lower_2.5,
    ymax = upper_97.5
  ), color = "#838383") +
    geom_point(colour = "#013369", size = 3) +
  coord_flip() +
  ylab("EPA") +
    theme_bw() +
  theme_light()+
  theme(plot.title = element_text(color="black", size=8, face="bold"))+
  theme(plot.title = element_text(size = 10, face = "bold"),
  plot.subtitle = element_text(size = 8))+
  theme(plot.background = element_rect(fill = "gray97"))+
  theme(panel.background = element_rect(fill = "gray97"),
        axis.title.y=element_blank(),
        axis.ticks.y=element_blank())+
  labs(title = "Individual Expected EPA Per Route Combo Against Cover 3: Top 15",
       subtitle = "How each route combo increases CP after controlling for coverage type, and QB | Min. 10 Attempts",
       caption = "Plot: Joseph Chernak, Data: SIS") +
  ggsave(filename = "Slide5_Bayes_EPA.png", 
       dpi = 300, type = "cairo", width = 10, height = 7, units = "in")
```


We can also view how an individual combo performs against each coverage.

```{r}
est  %>%
  filter(combo == "Go - Seam") %>%
  ggplot(aes(x = reorder(coverage, value), value)) +
  geom_linerange(size = .5, aes(
    ymin = lower_2.5,
    ymax = upper_97.5
  ), color = "#838383") +
    geom_point(colour = "#013369", size = 3) +
  coord_flip() +
  ylab("EPA") +
    theme_bw() +
  theme_light()+
  theme(plot.title = element_text(color="black", size=8, face="bold"))+
  theme(plot.title = element_text(size = 10, face = "bold"),
  plot.subtitle = element_text(size = 8))+
  theme(plot.background = element_rect(fill = "gray97"))+
  theme(panel.background = element_rect(fill = "gray97"),
        axis.title.y=element_blank(),
        axis.ticks.y=element_blank())+
  labs(title = "Individual EPA Estimate for the Go - Seam Combo Against Each Coverage",
       subtitle = "How the Go - Seam combo increases EPA after controlling for coverage type and QB (minimum 10 attempts)",
       caption = "Plot: Joseph Chernak, Data: SIS") +
  ggsave(filename = "Slide5_Go_Seam.png", 
       dpi = 300, type = "cairo", width = 10, height = 7, units = "in")
```

#save data
```{r}
write_rds(est, "EPA_Data.rds")
```





