---
title: "SIS Football Analytics Challenge 2021"
author: "Jeremy Siegel"
date: "7/7/2021"
output: html_document
---

# Which route combinations were most popular in the NFL in 2020. Of these combinations, which perform best against each coverage type

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Start by gathering our data
```{r}
library(tidyverse)
pbp <- read_csv(url("https://raw.githubusercontent.com/SportsInfoSolutions/AnalyticsChallenge2021/main/Data/PlayByPlay.csv"))

players <- read_csv(url("https://raw.githubusercontent.com/SportsInfoSolutions/AnalyticsChallenge2021/main/Data/SkillPositionPlayers.csv"))

```

Assumption: Now we are going to combine routes that are the same type but different direction (ex: Swing Left and Swing Right)
```{r}

players <- players %>% 
  mutate_at(vars(Route), ~case_when(
    startsWith(., "Swing") ~ "Swing",
    startsWith(., "Flat") ~ "Flat",
    startsWith(., "Screen") ~ "Screen",
    .  == "Chip" ~ "Blocking",
    startsWith(., "Chip") ~ sub("Chip - ", "", .),
    startsWith(., "Fade") ~ "Fade",
    TRUE ~ .
  )) %>% 
  filter(!(Route %in% c("NULL", "Blocking")))
```


First, let's take a look at the routes that are most commonly used in general
```{r}
route_counts <- players %>% count(Route) %>%  
arrange(desc(n))

route_counts
```
Now we will find out which routes are used on the most percentage of plays. This is a slightly different exercise than the last one, since each route will only be counted once if it is used on a given play. For example, if a given play has four different receivers running a slant, slant will only be counted once, rather than four times
```{r}
nPlays <- nrow(unique(players %>% select(GameID, EventID)))
players %>% select(GameID, EventID, Route) %>% 
  unique() %>% 
  group_by(Route) %>% 
  summarise(usage_pct = n()/nPlays) %>% 
  arrange(desc(usage_pct))
  
```

The results here are similar but slightly different from the previous list. Flat routes may not be used as often as out routes by they are used on a higher percentage of plays

Next we will divide each route into generic categories to see which family of routes are used most commonly, as identified by the route definitions on the SIS github

```{r}
short <- c("Curl", "Comeback", "Jerk", "Out", "Over Ball", "Whip")
vertical <- c("Go/Fly", "Fade", "Fade - Back Shoulder",
              "Seam", "Post", "Corner")
crossing <- c("Drag", "Dig", "Deep Cross", "Slant")
interior <- c("Angle", "Beneath", "Check & Release", "Chip",
              "Flat", "Leak", "Swing", "Wheel")
screen <- c("Beneath", "Bubble", "Quick", "Shovel", "Screen", "Jet Sweep Pass")

double_move <- c("Corner Post", "Post Corner", "Hitch & Go",
                 "Out & Up", "Sluggo", "Sitck-Nod")




players <- players %>% mutate(
 Family = case_when(
   Route %in% short ~ "short",
   Route %in% vertical ~ "vertical",
   Route %in% crossing ~ "crossing",
   Route %in% interior ~ "interior",
   Route %in% screen ~ "screen",
   Route %in% double_move ~ "double move",
   TRUE ~ "NULL"
 )) 

players %>% 
  filter(Family != "NULL") %>% 
  count(Family) %>% 
  arrange(desc(n))
```


Now we will find out how much the usage of each family of routes will increase (or decrease) the porbability of a completion against each coverage type
```{r}

pbp <- pbp %>% 
  filter(!CoverageScheme %in% c("Spike", "Other", "NULL", "Screen")) %>% 
   mutate_at(vars(Completion, OffensiveYardage, EPA, Attempt), as.numeric)
  

##Average completion pct against each coverage type

averages <- pbp %>% 
  group_by(CoverageScheme) %>% 
  summarise(comp_pct = mean(Completion, na.rm = T))

players <- players %>% filter(Family != "NULL")


pbp_all <- pbp %>% 
  inner_join(players, by = c("GameID", "EventID")) %>% 
  mutate_at(vars(Family, CoverageScheme), as.factor)

comp_mat <- matrix(NA, nrow = length(averages$CoverageScheme), ncol = length(levels(pbp_all$Family)),
                   dimnames = list(
                     levels(pbp_all$CoverageScheme),
                     levels(pbp_all$Family)
                   ))

```

This following matrix will represent each family's completion percentage against each route on a scale where 100 is average


```{r}
for(i in 1:length(averages$CoverageScheme)){
  for(j  in 1:length(levels(pbp_all$Family))){
    comp_mat[i,j] <- 100*(sum(pbp_all$Completion == 1 & pbp_all$CoverageScheme == averages$CoverageScheme[i] & 
                            pbp_all$Family == levels(pbp_all$Family)[j], na.rm = T)/ sum(pbp_all$CoverageScheme == averages$CoverageScheme[i] & 
                            pbp_all$Family == levels(pbp_all$Family)[j], na.rm = T))/
      as.numeric(averages[i,2])
      
  }
}
comp_mat
```

The most successful route in terms of predicting a completion is the screen against prevent coverage schemes, as it results in a completion 57 percent more than an average play against such defense. This makes sense, as the screen involves a quick, easy completion for the offense and prevent defense will allow short completions.

It seems as if the most extreme outcomes come from screen routes and prevent defenses. Let's see what happens if we use a different metric, such as yards per pass attempt

```{r}



averages <- pbp %>% 
  group_by(CoverageScheme) %>% 
  summarise(ypa = sum(OffensiveYardage * Attempt, na.rm =T)/
              sum(Attempt, na.rm = T))


for(i in 1:length(averages$CoverageScheme)){
  for(j  in 1:length(levels(pbp_all$Family))){
    comp_mat[i,j] <- 100*(sum(pbp_all$OffensiveYardage * pbp_all$Attempt * (pbp_all$CoverageScheme == averages$CoverageScheme[i]) *
                            (pbp_all$Family == levels(pbp_all$Family)[j]), na.rm = T)/ sum( pbp_all$Attempt * (pbp_all$CoverageScheme == averages$CoverageScheme[i]) *
                           (pbp_all$Family == levels(pbp_all$Family)[j]), na.rm = T))/
      as.numeric(averages[i,2])
      
  }
}
comp_mat



```
Here, we have slightly different results as the double move is super effective against Combo coverage schemes an Cover 0. Meanwhile, interior routes perform poorly against Cover 0.

Now let's try EPA per pass attempt
```{r}
averages <- pbp %>% 
  group_by(CoverageScheme) %>% 
  summarise(epa_pass = sum(EPA * Attempt, na.rm =T)/
              sum(Attempt, na.rm = T))


for(i in 1:length(averages$CoverageScheme)){
  for(j  in 1:length(levels(pbp_all$Family))){
    comp_mat[i,j] <- 100*(sum(pbp_all$EPA * pbp_all$Attempt * (pbp_all$CoverageScheme == averages$CoverageScheme[i]) *
                            (pbp_all$Family == levels(pbp_all$Family)[j]), na.rm = T)/ sum( pbp_all$Attempt * (pbp_all$CoverageScheme == averages$CoverageScheme[i]) *
                           (pbp_all$Family == levels(pbp_all$Family)[j]), na.rm = T))/
      as.numeric(averages[i,2])
      
  }
}
comp_mat

```
Here, we have similar, but more extreme results from the YPA trial. For example, a double move performs an insane 511 percent above average against Combo coverage schemes.  


## Final Answer

The most effective route family against each coverage scheme:

Combo: Double Move

Cover 0: Double Move

Cover 1: Crossing

Cover 2: Double  Move

Cover 3: Double Move

Cover 4: Double Move

Cover 6: Screen

Man Cover 2: Screen

Prevent: Screen

Tampa 2: Screen


The least effective route family against each coverage scheme:

Combo: Screen

Cover 0: Short

Cover 1: Short

Cover 2: Crossing

Cover 3: Short

Cover 4: Screen

Cover 6: Double Move

Man Cover 2: Double Move

Prevent: Interior

Tampa 2: Crossing


